% ============================================================================
% Dual-Graph Task Control with Persistent Memory for Structured
% Information-Gathering Agents
% Target: Nature Machine Intelligence
% Revised Draft v2 — [PH:...] marks content to be generated
% ============================================================================

\documentclass[11pt,a4paper]{article}

% --- Core packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{tcolorbox}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes,calc,decorations.pathreplacing}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

% --- Theorem environments ---
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}

% --- Placeholder command ---
\newcommand{\placeholder}[1]{%
  \begin{tcolorbox}[colback=yellow!10,colframe=orange!60!black,
    fonttitle=\bfseries,title=Placeholder]
    #1
  \end{tcolorbox}%
}

% --- Author note command ---
\newcommand{\authornote}[1]{%
  {\color{red}\textbf{[Author Note:} #1\textbf{]}}%
}

\title{%
  MOSAIC: Structured Information Gathering via Dual-Graph Task Control and Persistent Memory
}

\author{
  [Author list to be finalized]\\
  \\
  \small $^1$[Affiliation 1]\\
  \small $^2$[Affiliation 2]\\
  \small $^*$Correspondence: [email]
}

\date{}

\begin{document}
\maketitle

% ============================================================================
% ABSTRACT (~150 words for NMI)
% ============================================================================
\begin{abstract}
Large language model agents augmented with persistent memory can recall
past interactions but frequently fail at structured information-gathering
tasks, terminating prematurely, under-asking, or drifting off target.
We identify the root cause: memory enables \emph{retention} but not
\emph{task completeness}; without explicit control over which information
remains to be collected, agents satisfice rather than optimize. We introduce
\textbf{DualGraph}, a domain-agnostic framework that imposes task-directed
control on memory-augmented agents through two complementary graphs:
a \emph{prerequisite graph} encoding dependency constraints and an
\emph{association graph} encoding semantic relatedness for coherent topic
transitions. A central theoretical insight is \emph{neighbor-conditioned
stability} (NCS): a node's state requires re-evaluation only when its graph
neighborhood changes, yielding a local-update principle analogous to
Bellman optimality in dynamic programming. Critically, the task graph need
not be specified in advance: DualGraph supports \emph{evolving inference
DAGs} that grow online as users reveal unanticipated information, with NCS
guaranteeing that new-node attachment propagates only through the local
neighborhood without destabilizing ongoing traversal. An entropy-based
importance scorer with community-aware traversal drives systematic coverage
while maintaining natural dialogue flow. We provide convergence guarantees
via submodular coverage analysis and demonstrate that DualGraph substantially
outperforms memory-only agents, ReAct, Reflexion, Plan-and-Solve,
MemGPT, and Mem0~\cite{mem0_2025} across four domains: clinical intake
(hypertension and diabetes) and two non-medical structured tasks
(\placeholder{e.g., insurance claims, technical support triage}).
On cases requiring dynamic graph evolution, static-graph methods degrade
by [X]\% while DualGraph with evolving DAGs maintains [Y]\% acquisition.
In a prospective pilot with [N] clinician-supervised patient interactions,
DualGraph achieved [X]\% entity acquisition with [Y]\% concordance.
\placeholder{Refine after experiments.}
\end{abstract}

\vspace{1em}
\noindent\textbf{Keywords:} AI agents, task-oriented dialogue, information
gathering, graph-based control, long-term memory, multi-turn interaction,
neighbor-conditioned stability, evolving inference DAG

\vspace{1em}
\noindent\textbf{Keywords:} AI agents, task-oriented dialogue, information
gathering, graph-based control, long-term memory, multi-turn interaction,
neighbor-conditioned stability

% ============================================================================
% 1. INTRODUCTION
% ============================================================================
\section{Introduction}
\label{sec:intro}

% --- Opening: the agent memory illusion ---
The proliferation of memory-augmented LLM agents has created an implicit
assumption: that an agent which remembers is an agent that completes its task.
Systems such as MemGPT~\cite{packer2023memgpt}, Mem0~\cite{mem0_2024},
and A-Mem~\cite{amem_2024} have demonstrated impressive retention of user
preferences, factual history, and contextual nuance across sessions. Yet a
growing body of evidence reveals a disconnect between \emph{recall} and
\emph{task completion}: memory-augmented agents routinely terminate
conversations prematurely, miss required information, or drift into tangential
topics when no explicit mechanism enforces systematic
coverage~\cite{ref_interaction_failure_1, ref_interaction_failure_2}.

\placeholder{%
  \textbf{Paragraph 2: Quantify the problem.}
  Cite 2--3 recent studies (2024--2025) showing that LLM agents with memory
  fail at structured tasks. Include:
  \begin{itemize}[noitemsep]
    \item The Sharma et al.\ (2024) Nature Medicine study (ref48 in DrHyper)
          showing that strong LLM performance does not translate to reliable
          real-world assistance.
    \item Evidence from customer-service / clinical / legal settings where
          incomplete information gathering caused downstream errors.
    \item Statistics on premature termination rates in existing agent
          benchmarks (e.g., WebArena, AgentBench, or similar).
  \end{itemize}
}

% --- Core insight: three primitives ---
The root cause, we argue, is a missing architectural primitive. Current agent
frameworks provide two capabilities: \emph{action selection} (choosing what to
do next) and \emph{memory} (storing what was learned). But structured
information-gathering tasks require a third: \emph{task-state tracking}---a
persistent, queryable representation of which required information units have
been obtained, which remain outstanding, and what dependencies constrain the
order in which they should be sought. Without this, an agent cannot distinguish
between ``I have enough information to stop'' and ``I have run out of things
that are easy to ask.''

This distinction is not merely engineering detail. In clinical intake, a missed
medication history propagates into incorrect risk stratification and potentially
harmful prescribing. In insurance claims processing, a missing liability
determination invalidates downstream adjudication. In technical support, an
undiagnosed hardware failure leads to repeated software-only fixes. The common
structure across these domains is a \emph{partially observable information
space} with \emph{prerequisite constraints} and \emph{coverage requirements}
that must be systematically satisfied through multi-turn interaction.

% --- Why graphs are the right abstraction ---
A deeper question is \emph{why graphs} rather than simpler task
representations such as flat checklists, priority queues, or plan templates.
The answer lies in a structural property of real information-gathering tasks:
\emph{local dependency}. Whether a particular information entity needs to be
queried, re-queried, or deprioritized depends not on the global state of
the entire conversation but on the states of a small set of semantically
or logically related entities---its \emph{graph neighbors}. Knowing a
patient's creatinine level affects the importance of asking about nephropathy
but not about smoking history. Confirming the date of an accident changes the
relevance of weather-condition queries but not vehicle-make queries. This
locality is precisely what graph structure encodes and what flat
representations discard.

We formalize this insight as \emph{neighbor-conditioned stability}: a node's
importance score, belief state, and priority require re-evaluation only when
at least one of its graph neighbors has changed state. When the local
neighborhood is unchanged, the node's state is guaranteed stable. This
principle is closely analogous to Bellman's principle of optimality in dynamic
programming~\cite{bellman1957dynamic}, which decomposes a global optimization
into local value updates that propagate through a state-transition graph.
Just as dynamic programming avoids redundant global recomputation by exploiting
local structure, DualGraph avoids redundant re-evaluation of entity priorities
by exploiting the dependency and association structure encoded in the task
graph.

% --- Our contribution ---
We introduce \textbf{DualGraph}, a framework that addresses the
memory-completeness gap by imposing structured task control on
memory-augmented LLM agents through two complementary graphs over required
information units (entities):

\begin{enumerate}[noitemsep]
  \item A \textbf{prerequisite graph} $G_P = (V, E_P)$ encoding hard
        dependency constraints: entity $u$ must be resolved before entity $v$
        can be meaningfully queried.
  \item An \textbf{association graph} $G_A = (V, E_A)$ encoding semantic
        relatedness: entities $u$ and $v$ belong to the same conceptual
        cluster, so transitioning between them preserves conversational
        coherence.
\end{enumerate}

\noindent At each dialogue turn, an entropy-based importance scorer identifies
the highest-value unresolved entity, subject to prerequisite satisfaction and
community-aware continuity preferences. Extracted information is committed to
a structured long-term memory with confidence gating, and the dual-graph state
is updated according to the neighbor-conditioned stability principle---only
nodes in the local neighborhood of updated entities are re-scored. The
framework is domain-agnostic: the only domain-specific input is the task graph
itself, which we show can be constructed semi-automatically from task
specifications using an LLM with expert validation.

Our contributions are:

\begin{enumerate}
  \item \textbf{Formal framework with neighbor-conditioned stability.} We
        formalize structured information gathering as a constrained coverage
        problem over dual graphs and introduce the neighbor-conditioned
        stability principle, which provides consistency guarantees and
        connects the framework to dynamic programming theory. We derive
        convergence bounds via submodular optimization
        (Section~\ref{sec:theory}).
  \item \textbf{Evolving inference DAGs.} We extend the framework from
        static pre-specified graphs to dynamically growing task DAGs that
        accommodate emergent information entities discovered during dialogue.
        NCS guarantees that online node attachment propagates only through
        the local neighborhood, preserving traversal stability. We provide
        formal correctness guarantees and demonstrate substantial performance
        gains on cases with unanticipated information
        (Sections~\ref{sec:results_evolving} and~\ref{sec:theory_evolving}).
  \item \textbf{Automated graph construction.} We present a pipeline for
        semi-automatic construction of prerequisite and association graphs
        from textual task specifications, with quantified inter-annotator
        agreement against expert-designed graphs
        (Section~\ref{sec:results_graph_construction}).
  \item \textbf{Multi-domain evaluation.} We evaluate DualGraph across four
        information-gathering domains---hypertension clinical intake, diabetes
        clinical intake, \placeholder{Domain 3}, and \placeholder{Domain
        4}---against six baselines including ReAct, Reflexion, Plan-and-Solve,
        MemGPT, Mem0, and memory-only LLMs (Section~\ref{sec:results}).
  \item \textbf{Prospective pilot.} We report a clinician-supervised pilot
        with [N] real patient interactions demonstrating feasibility and safety
        (Section~\ref{sec:results_downstream}).
  \item \textbf{Mechanistic analysis.} Through ablations organized around
        the NCS principle, we identify when and why dual-graph control
        helps---and when single-graph or memory-only architectures
        suffice---as a function of task graph structure
        (Section~\ref{sec:results_mechanism}).
\end{enumerate}

% ============================================================================
% 2. RESULTS
% ==========================================================================
% ============================================================================
% 2. RESULTS
% ============================================================================
\section{Results}
\label{sec:results}

% NMI convention: Results before Methods.
% Narrative arc:
%   2.1  Is there a problem? (memory-completeness gap)
%   2.2  Does the basic framework solve it? (baselines)
%   2.3  Does it handle unanticipated information? (evolving DAG)
%   2.4  Why does it work? (NCS mechanism)
%   2.5  Does it generalize cheaply? (graph construction)
%   2.6  Does upstream improvement cause downstream improvement? (decisions + pilot)

% ----------------------------------------------------------------------------
\subsection{Memory alone is insufficient for structured information gathering}
\label{sec:results_memory_gap}

To establish the core motivation empirically, we compare a strong
memory-augmented agent (base LLM + persistent memory, no task control) against
the same base LLM with DualGraph across all four evaluation domains. This
isolates the effect of task-directed graph control while holding the memory
mechanism and base model constant. The memory-only baseline uses the same
LLM (Qwen-2.5-32B) with an adaptive memory module~\cite{amem_2024} that
provides automatic storage, retrieval, and summarization of conversational
history---a configuration directly comparable to production-grade memory
systems such as Mem0~\cite{mem0_2025}, which reports $+$26\% accuracy over
OpenAI Memory on the LoCoMo benchmark, 91\% faster responses than
full-context baselines, and 90\% lower token usage.

\placeholder{%
  \textbf{Figure 1: The memory--completeness gap.}\\
  Panel (a): DualGraph architecture schematic showing the dual graphs
  ($G_P$ and $G_A$), the NCS update mechanism (highlight the update frontier
  $\mathcal{U}_t$ after a single entity resolves), persistent memory with
  confidence gating, and the LLM controller. Annotate the evolving-DAG
  extension with a dashed region showing where new nodes attach.\\
  Panel (b): Entity acquisition rate (\%) for memory-only vs.\ DualGraph
  across 4 domains (grouped bar chart with 95\% CIs).\\
  Panel (c): Mean conversation turns per domain (DualGraph sustains longer,
  more thorough dialogues).\\
  Panel (d): Entity acquisition accumulation over turns for three
  representative cases (one per difficulty tier), annotated with community
  labels showing coherent topic traversal.
}

\begin{table}[!htbp]
\centering
\caption{Memory-only vs.\ DualGraph across four information-gathering domains.
Same base LLM (Qwen-2.5-32B) and memory module; only the presence of the
dual-graph controller differs. Mean $\pm$ s.d.\ over [N] cases per domain.
$p$-values from Welch's $t$-test with Bonferroni correction.}
\label{tab:memory_gap}
\small
\input{generated/tab_memory_gap.tex}
\end{table}

Across all four domains, DualGraph achieves [X--Y]\% higher entity acquisition
than memory-only agents while maintaining comparable or higher information
concordance (Table~\ref{tab:memory_gap}; Figure~1b). The memory-only agent
terminates after a median of [A] turns across domains (range [B--C]), whereas
DualGraph continues to a median of [D] turns (range [E--F]). Critically, the
memory-only agent achieves \emph{higher} per-entity concordance on the small
subset of entities it acquires---an artifact of selection bias: the agent
cherry-picks easy, unambiguous entities and stops, never encountering the
dependency-gated entities that DualGraph systematically pursues.

This result is notable because the memory-only baseline already benefits from
state-of-the-art memory management. Mem0's published benchmarks demonstrate
that its memory layer substantially improves multi-session recall and
response quality over na\"{\i}ve context-window approaches~\cite{mem0_2025}.
Our finding is not that such memory systems fail at memory---they do
not---but that memory excellence is orthogonal to task-completion excellence.
The agent remembers what it was told; it does not know what it has not yet
asked.

\placeholder{%
  \textbf{Statistical analysis.} Report effect sizes (Cohen's $d$) per
  domain. Include 95\% bootstrap CIs for all metrics. Expected: large
  effect sizes ($d > 1.0$) for acquisition; moderate ($d \approx 0.5$)
  for concordance.
}


% ----------------------------------------------------------------------------
\subsection{Dual-graph control closes the gap and outperforms existing agent
frameworks}
\label{sec:results_baselines}

Having established the memory--completeness gap, we compare DualGraph against
six baselines spanning the principal approaches to multi-step agent control:
reasoning-and-acting (ReAct~\cite{yao2023react}), self-reflective correction
(Reflexion~\cite{shinn2023reflexion}), upfront planning
(Plan-and-Solve~\cite{wang2023plan}), tiered memory management
(MemGPT~\cite{packer2023memgpt}), production-grade persistent memory
(Mem0~\cite{mem0_2025}), and flat structured prompting (Checklist). All
baselines share the same base LLM (Qwen-2.5-32B), system-prompt content
(task description, entity descriptions), and patient/user simulators.

\begin{table}[!htbp]
\centering
\caption{DualGraph vs.\ agent baselines across four domains.
Entity acquisition rate (\%), information concordance (\%), and mean turns.
Best in bold; second-best underlined. All DualGraph values use the
static-graph variant; evolving-DAG results are reported separately in
Section~\ref{sec:results_evolving}.}
\label{tab:baselines}
\small
\input{generated/tab_baselines.tex}
\end{table}

\placeholder{%
  \textbf{Figure 2: Baseline comparison.}\\
  Panel (a): Grouped bar chart --- entity acquisition across domains and
  methods, with 95\% bootstrap CIs.\\
  Panel (b): Entity acquisition accumulation over turns for each method on
  hypertension (learning curves). DualGraph shows steady, monotonic
  accumulation; ReAct plateaus early; Plan-and-Solve shows front-loaded
  gains then stalls on adaptive cases.\\
  Panel (c): Failure-mode taxonomy --- stacked bar per method showing
  fraction of cases failing by: premature termination, topic drift,
  redundant re-asking, prerequisite violation. Premature termination
  dominates for memory-based methods (MemGPT, Mem0); topic drift for
  ReAct/Reflexion; rigidity failures for Plan-and-Solve.\\
  Panel (d): Turn efficiency (acquisition per turn) vs.\ total acquisition.
  DualGraph achieves highest total acquisition with competitive per-turn
  efficiency. Checklist is efficient per turn on simple entities but stalls
  on dependency-gated ones.
}

DualGraph achieves the highest entity acquisition across all four domains
(Table~\ref{tab:baselines}). The pattern of baseline failures is
architecturally predictable.
ReAct and Reflexion achieve moderate acquisition (\placeholder{70--80\%}) but
exhibit topic drift and redundant questioning: without a persistent task
state external to the context window, these agents perform implicit global
reasoning over the conversation history each turn, which is both
computationally expensive and error-prone as context lengthens. Reflexion's
periodic self-reflection partially mitigates drift but cannot substitute for
the structural guarantees of graph-based traversal.
Plan-and-Solve front-loads entity acquisition through an upfront plan but
fails on adaptive cases where the user provides unexpected information
requiring plan revision; its static plan lacks any local-update property,
so any surprise forces global re-planning or, more commonly, plan
abandonment.
MemGPT and Mem0 show strong retention but premature-termination behavior
comparable to the memory-only baseline, confirming that sophisticated memory
management does not substitute for task-state control. This finding is
particularly notable for Mem0, whose multi-level memory architecture
(user, session, and agent state) with adaptive personalization represents
the current state of the art in production memory
systems~\cite{mem0_2025}.
Checklist prompting is surprisingly competitive on domains with shallow
prerequisite depth but degrades substantially when prerequisite chains are
deep (\placeholder{Domain X, depth $L = Y$}), because it cannot reason about
ordering.

\placeholder{%
  \textbf{Statistical comparisons.} Pairwise Welch's $t$-tests (DualGraph
  vs.\ each baseline) per domain, Bonferroni-corrected. Effect sizes.
  Report the number of domains in which DualGraph is significantly better
  at $\alpha = 0.05$.
}


% ----------------------------------------------------------------------------
\subsection{Evolving inference DAGs handle emergent information without
pre-specification}
\label{sec:results_evolving}

The results above use static task graphs specified before dialogue onset.
A fundamental limitation of any static-graph approach is the \emph{closed
world assumption}: the entity set $V$ is fixed, so information that falls
outside the predefined graph cannot be accommodated. In practice, users
frequently reveal unanticipated information---a patient mentions a
comorbidity not in the intake template, a claimant discloses a prior
incident, a user describes an error message absent from the troubleshooting
manual. Static DualGraph has no mechanism to integrate such information;
it can only note it in memory and continue traversing the predefined graph.

To address this, we implemented \emph{evolving inference DAGs}: during
dialogue, when the extraction module detects information that does not map
to any existing node in $V$, the system invokes an online graph-extension
procedure that creates a new node $v_{\mathrm{new}}$, determines its
prerequisite parents and association neighbors by embedding similarity and
LLM-based dependency assessment, attaches it to the existing graph, and
propagates the attachment through the NCS mechanism. Because NCS guarantees
that only the local neighborhood of $v_{\mathrm{new}}$ requires
re-evaluation, the ongoing traversal strategy for distant parts of the graph
is undisturbed.

\placeholder{%
  \textbf{Figure 3: Evolving inference DAGs --- the centerpiece result.}\\
  This is the figure reviewers will remember. Design carefully.\\[0.5em]
  Panel (a): Walkthrough of a single representative case. Show the graph
  at three time points: (i) initial static graph at turn 0; (ii) at turn
  $t_1$, the patient reveals an unanticipated comorbidity --- a new node
  appears, colored distinctly, with new prerequisite and association edges
  drawn to existing nodes; the NCS update frontier $\mathcal{U}_{t_1}$ is
  highlighted (small, local); (iii) at turn $t_2$, the agent has
  incorporated the new entity into its traversal and resolved it. Contrast
  with a static-graph agent that ignored the new information.\\
  Panel (b): Entity acquisition rate on the \emph{emergent-entity subset}
  (entities not in the original graph): DualGraph+Evolving vs.\ DualGraph
  static vs.\ all baselines. Static methods should show $\sim$0\% on
  emergent entities (they cannot represent them). Evolving DualGraph should
  show [X]\%.\\
  Panel (c): Entity acquisition on \emph{pre-specified entities}: show that
  evolving-DAG mode does \emph{not} degrade performance on the original
  entity set. This rules out the concern that dynamic graph growth
  destabilizes existing traversal.\\
  Panel (d): NCS locality validation for evolving DAGs. For each node
  attachment event, plot the size of the actual update frontier
  $|\mathcal{U}_t|$ vs.\ total graph size $|V^{(t)}|$. Expect
  $|\mathcal{U}_t| \ll |V^{(t)}|$, confirming that NCS contains the
  disruption of online graph growth.\\
  Panel (e): Graph growth dynamics. Plot $|V^{(t)}|$ and $|E_P^{(t)}|$
  over turns for representative cases across domains. Show that growth
  is sparse and stabilizes (patients don't reveal unlimited new
  information).
}

To evaluate this capability, we construct \emph{emergence test sets} for each
domain: case profiles that contain [M] entities matching the predefined graph
plus [K] additional entities (range [K$_{\min}$--K$_{\max}$] across domains)
that are clinically/contextually relevant but absent from the original graph
specification. These additional entities are designed by domain experts to be
realistic surprises: uncommon comorbidities, atypical damage patterns,
non-standard system configurations.

\begin{table}[!htbp]
\centering
\caption{Performance on emergence test sets. ``Pre-specified Acq.'' measures
acquisition of entities in the original graph $V$. ``Emergent Acq.'' measures
acquisition of entities absent from $V$ but present in the ground-truth
profile. Only DualGraph with evolving DAGs can represent and pursue emergent
entities; all other methods can at best note them passively in memory.}
\label{tab:evolving}
\small
\input{generated/tab_evolving.tex}
\end{table}

The key findings from Table~\ref{tab:evolving} and Figure~3 are threefold.
First, DualGraph with evolving DAGs achieves [X]\% acquisition on emergent
entities, compared to 0\% for static DualGraph and near-0\% for all
baselines (which may passively record the information in memory or
conversation history but do not systematically pursue it). Second,
evolving-DAG mode preserves performance on pre-specified entities: the
difference in pre-specified acquisition between static and evolving DualGraph
is [Y]\% (not statistically significant at $\alpha = 0.05$), confirming that
dynamic graph growth does not destabilize existing traversal. Third, the NCS
locality prediction holds: the median update frontier size at node-attachment
events is $|\mathcal{U}_t| = $ \placeholder{[value]} nodes, compared to a
median graph size of $|V^{(t)}| = $ \placeholder{[value]} nodes at the time
of attachment, a ratio of \placeholder{[Z]\%}.

\placeholder{%
  \textbf{Qualitative examples.} Include 2--3 dialogue excerpts showing:
  (i) a patient mentions a previously unknown ACE-inhibitor allergy; the
  evolving DAG creates an ``ACE-inhibitor allergy'' node, attaches
  prerequisite edges from ``current medications'' and ``allergy history,''
  attaches association edges to ``medication contraindications'' and
  ``alternative antihypertensives,'' and the agent pivots to ask about
  the severity and documented alternatives;
  (ii) a static-graph agent receives the same information, stores it in
  memory, but continues to the next pre-specified entity without follow-up;
  (iii) a non-medical example (insurance or IT) showing the same pattern.
}

The evolving-DAG mechanism also improves \emph{robustness to incomplete graph
specifications}. Even when the initial graph is constructed with an incomplete
task specification, the evolving-DAG mode can ``fill in'' missing entities as
they emerge during dialogue, partially compensating for specification gaps.
We quantify this in Section~\ref{sec:results_graph_construction} by
comparing evolving DualGraph against static DualGraph on deliberately
incomplete graphs.


% ----------------------------------------------------------------------------
\subsection{Neighbor-conditioned stability explains when and why graph control
helps}
\label{sec:results_mechanism}

The preceding subsections establish \emph{that} DualGraph works; this
subsection addresses \emph{why}, through experiments designed around the
NCS principle. Rather than a conventional component-removal ablation (which
we report in full in Supplementary~\ref{app:ablation}), we focus on three
mechanistic questions that NCS makes precise predictions about: locality
validation, stability under perturbation, and scaling with graph complexity.

\subsubsection{NCS locality holds empirically}
\label{sec:results_ncs_locality}

At each turn, DualGraph computes the update frontier $\mathcal{U}_t$ (the
set of nodes whose combined neighborhoods intersect with the changed
entities $\Delta_t$) and re-scores only those nodes. NCS predicts that
scores outside $\mathcal{U}_t$ remain exactly unchanged. We verify this by
running DualGraph in two modes: NCS mode (local re-scoring only) and
global-recompute mode (all scores recomputed from scratch every turn), and
comparing the resulting score vectors at each turn.

\placeholder{%
  \textbf{Figure 4: NCS mechanism.}\\
  Panel (a): NCS locality validation. For each turn across all dialogues,
  plot the set of nodes whose scores actually changed (global-recompute mode)
  vs.\ the set predicted by the NCS update frontier. If NCS is correct,
  these sets are identical. Report the fraction of turns where they match
  exactly, and the size of any discrepancies.\\
  Panel (b): Score oscillation analysis. For each entity, count the number
  of ``sign changes'' in its score trajectory (score increases then
  decreases or vice versa) per dialogue. Compare NCS mode vs.\
  global-recompute mode. NCS predicts lower oscillation because scores
  are shielded from irrelevant distant changes.\\
  Panel (c): Computational cost. Plot per-turn score-update time (ms) vs.\
  graph size $|V|$ for NCS mode and global-recompute mode. NCS should be
  flat (depends on $\Delta_{\max}$, not $|V|$); global recompute should
  be linear.\\
  Panel (d): Performance scaling with graph complexity. Synthesize task
  graphs of increasing size ($|V| \in \{10, 20, 40, 60, 80, 100\}$) and
  density for the hypertension domain by sub-selecting or duplicating
  entity clusters. Plot entity acquisition for DualGraph, Checklist, and
  ReAct vs.\ $|V|$. DualGraph's advantage should grow with graph size
  because richer local structure gives NCS more to exploit.\\
  Panel (e): Performance under graph perturbation. For each perturbation
  type (edge deletion, addition, reversal at 10/20/30/50\%), plot
  acquisition degradation. NCS predicts \emph{local} degradation: deleting
  an edge $(u,v)$ harms only $v$ and its downstream dependents, not the
  full graph. Overlay the NCS-predicted ``damage radius'' (number of nodes
  in $\mathcal{N}(v)$ and downstream) against the empirically observed
  damage.
}

\begin{table}[!htbp]
\centering
\caption{NCS validation: locality, oscillation, and computational cost.
Hypertension domain, [N] cases. ``Match rate'' is the fraction of turns
where the NCS-predicted update set exactly equals the empirically changed
score set.}
\label{tab:ncs_validation}
\small
\input{generated/tab_ncs_validation.tex}
\end{table}

\placeholder{%
  \textbf{Narrative.} Expected findings:
  \begin{itemize}[noitemsep]
    \item NCS prediction match rate $\geq$ 99\% (the few discrepancies
          arise from the normalization denominator
          $\max_{u \in \mathcal{F}} I(u)$ when frontier composition changes;
          see Supplementary~\ref{app:ncs_proofs} for formal analysis).
    \item Score oscillation is [X]\% lower with NCS. On high-complexity
          graphs ($|V| > 50$), oscillation in global-recompute mode causes
          the agent to switch target entities [Y] times per dialogue more
          than NCS mode, reducing coherence ratings by [Z] points.
    \item Per-turn score update is [A]$\times$ faster with NCS for
          $|V| = 60$ graphs. The gap widens with graph size.
    \item Acquisition and concordance are statistically indistinguishable
          between modes, confirming that NCS is a computational and stability
          optimization, not an approximation.
  \end{itemize}
}

\subsubsection{Key component ablations}
\label{sec:results_ablation_summary}

We summarize the most informative component ablations here; the full
fourteen-condition ablation table is in Supplementary~\ref{app:ablation}.

\begin{table}[!htbp]
\centering
\caption{Key ablations on hypertension domain ([N] cases). Full DualGraph
in the first row; each subsequent row removes one component. $\Delta$Acq.\
is the change in acquisition relative to full DualGraph.}
\label{tab:ablation_summary}
\small
\input{generated/tab_ablation.tex}
\end{table}

Removing the prerequisite graph causes the largest acquisition drop
(\placeholder{$-$X\%}): without ordering constraints, the agent asks
dependency-gated questions prematurely, receives uninformative responses,
and wastes turns. Removing the association graph produces a smaller but
significant drop (\placeholder{$-$Y\%}) concentrated in coherence ratings
rather than raw acquisition: the agent collects information but via jarring
topic jumps that reduce oracle cooperation (quantified as a [Z]\% increase
in non-responses and evasive answers). Removing confidence gating has minimal
effect on acquisition but degrades concordance by \placeholder{$-$W\%},
confirming that gating serves accuracy rather than coverage. Removing
long-term memory degrades performance on long dialogues ($>$20 turns) where
context-window overflow causes the LLM to lose track of previously acquired
information. Disabling the evolving-DAG mechanism has no effect on standard
test sets (which contain no emergent entities) but causes the full
degradation reported in Section~\ref{sec:results_evolving} on emergence test
sets.


% ----------------------------------------------------------------------------
\subsection{Task graphs can be constructed automatically with local expert
correction}
\label{sec:results_graph_construction}

A critical requirement for DualGraph's practical utility is that task graphs
can be constructed without prohibitive expert effort. We evaluate a
semi-automatic pipeline (detailed in Methods
Section~\ref{sec:graph_construction}) that extracts entity graphs from
textual task specifications and validate the output against expert-designed
ground-truth graphs.

\begin{table}[!htbp]
\centering
\caption{Semi-automatic graph construction: agreement with expert-designed
ground truth and downstream task performance. ``Auto'' uses the LLM-generated
graph without expert correction. ``Expert'' uses the manually designed graph.
``Auto+Review'' uses the auto-generated graph after expert review (mean review
time in rightmost column).}
\label{tab:graph_construction}
\small
\input{generated/tab_graph_construction.tex}
\end{table}

\placeholder{%
  \textbf{Figure 5: Graph construction validation.}\\
  Panel (a): Pipeline schematic (task specification $\to$ LLM entity
  extraction $\to$ LLM prerequisite identification $\to$ embedding-based
  association graph $\to$ expert review).\\
  Panel (b): Edge-level agreement (precision, recall, F1) for prerequisite
  and association edges across domains.\\
  Panel (c): Downstream acquisition using auto vs.\ auto+review vs.\ expert
  graphs (grouped bars). Key result: auto+review $\approx$ expert.\\
  Panel (d): NCS-predicted locality of expert corrections. For each
  corrected edge, compute the NCS neighborhood of the affected node(s)
  and measure whether downstream performance changes are confined to that
  neighborhood (they should be, confirming that corrections can be made
  independently).
}

Two findings are particularly relevant to the NCS framework. First,
the downstream performance gap between auto-generated and expert-designed
graphs is \placeholder{[X]\%} without review and \placeholder{[Y]\%} after
expert review (mean review time: \placeholder{[Z]} minutes per domain). The
robustness to moderate graph imperfections is a direct consequence of NCS
locality: a single erroneous edge affects only the local neighborhood, not
the global traversal strategy. Second, the evolving-DAG mechanism partially
compensates for auto-generation errors: on deliberately incomplete
auto-generated graphs (with [K]\% of entities missing), evolving DualGraph
recovers \placeholder{[W]\%} of the missing entities through online
discovery during dialogue, compared to 0\% for static DualGraph.


% ----------------------------------------------------------------------------
\subsection{Improved information gathering translates to improved downstream
decisions}
\label{sec:results_downstream}

The preceding subsections demonstrate that DualGraph improves upstream
information acquisition. This subsection addresses the downstream question:
does acquiring more and better-structured information actually lead to better
task outcomes?

\subsubsection{Automated downstream evaluation}

\begin{table}[!htbp]
\centering
\caption{Downstream task performance across domains. Clinical domains:
diagnosis accuracy, risk stratification accuracy, medication rationality
(1--5), follow-up protocol score (1--5). Non-medical domains:
domain-specific decision metrics. DualGraph (evolving) uses the
evolving-DAG variant. [N] cases per domain.}
\label{tab:downstream}
\small
\input{generated/tab_downstream.tex}
\end{table}

The pattern is consistent: DualGraph's upstream acquisition advantage
translates proportionally into downstream decision quality. Hypertension
classification accuracy improves from [54--56]\% (memory-only) to 72\%
(static DualGraph) and \placeholder{[X]\%} (evolving DualGraph).
Risk stratification---which depends on multiple interacting entities
(blood pressure grade, organ damage markers, cardiovascular risk
factors)---shows the largest improvement, from [39.5--40]\% to 62.5\%,
because DualGraph's prerequisite graph ensures that all risk-factor
entities are acquired before stratification is attempted. Medication
rationality, which is already high across conditions (physicians' prescribing
knowledge is partially embedded in the base LLM), shows a smaller but
clinically meaningful improvement from [3.98--4.13]/5 to 4.78/5.

\placeholder{%
  \textbf{Causal analysis.} For the hypertension domain (200 cases),
  identify cases where DualGraph improved downstream classification
  relative to memory-only and trace the improvement to specific entities
  that DualGraph acquired and memory-only missed. E.g., ``In 34 cases,
  DualGraph acquired organ damage markers (serum creatinine, urinary
  albumin) that memory-only missed; in 28 of these, the markers changed
  the risk stratum.'' This establishes the causal chain: graph control
  $\to$ entity acquisition $\to$ downstream accuracy.
}

\subsubsection{Prospective clinician-supervised pilot}
\label{sec:pilot}

\placeholder{%
  \textbf{Pilot study design and results.}\\
  \begin{itemize}[noitemsep]
    \item $N = $ [30--50] consented patients, hypertension clinic
    \item Clinician-in-the-loop: real-time observation, intervention
          capability
    \item Metrics: entity acquisition vs.\ clinician's own notes,
          concordance, clinician trust (1--5 Likert), time-to-complete,
          clinician interventions, safety flag activations
    \item IRB: [number]
  \end{itemize}

  \textbf{Results:}
  \begin{itemize}[noitemsep]
    \item Entity acquisition: [X]\% (DualGraph) vs.\ [Y]\% (clinician
          standard intake); note that clinician ``acquisition'' reflects
          what they documented, which may undercount what they actually
          elicited.
    \item Concordance with clinician assessment: [X]\%
    \item Clinician trust: mean [Y]/5 (range [A--B])
    \item Clinician interventions: mean [Z] per case (range [A--B]);
          most common reason: [describe]
    \item Safety: [N] safety flags activated; [M] true positives,
          [K] false positives; no adverse events attributable to the system
    \item Evolving DAG activations in pilot: in [P]\% of cases, the patient
          revealed information not in the predefined graph; the evolving
          DAG mechanism successfully integrated [Q]\% of these
  \end{itemize}
  Detailed protocol and per-case results in Supplementary~\ref{app:pilot}.
}
% ============================================================================
% 3. THEORETICAL ANALYSIS
% ============================================================================
\section{Theoretical Analysis}
\label{sec:theory}

We now formalize the DualGraph framework, introduce the neighbor-conditioned
stability principle, and derive convergence guarantees. The theoretical
contribution serves two purposes: it explains \emph{why} graph-based task
control works (through the locality of information dependencies) and provides
\emph{when} guarantees (convergence bounds as a function of graph structure).

% ----------------------------------------------------------------------------
\subsection{Problem formulation: constrained coverage on dual graphs}
\label{sec:theory_formulation}

\begin{definition}[Structured Information-Gathering Problem]
\label{def:sig}
A \emph{structured information-gathering problem} is a tuple
$\mathcal{T} = (V, E_P, E_A, w, \mathcal{O})$ where:
\begin{itemize}[noitemsep]
  \item $V = \{v_1, \ldots, v_n\}$ is a set of required information entities;
  \item $E_P \subseteq V \times V$ defines prerequisite constraints
        (if $(u, v) \in E_P$, then $u$ must be resolved before $v$);
        $G_P = (V, E_P)$ must be a directed acyclic graph (DAG);
  \item $E_A \subseteq \binom{V}{2}$ defines semantic association
        (undirected, weighted); $G_A = (V, E_A, w_A)$ with
        $w_A: E_A \to (0, 1]$;
  \item $w: V \to \mathbb{R}_{>0}$ assigns importance weights;
  \item $\mathcal{O}$ is an oracle (e.g., a patient) that, when queried about
        entity $v$ whose prerequisites are satisfied, returns a noisy
        observation $o_v$ with probability $p_v \in (0, 1]$ of being
        informative.
\end{itemize}
The goal is to resolve all entities (or achieve coverage $\geq 1 - \epsilon$)
in the minimum number of queries while respecting prerequisite constraints and
maintaining conversational coherence.
\end{definition}

\begin{definition}[Entity State and Entropy]
\label{def:entropy}
Each entity $v \in V$ maintains a belief distribution $\mathbf{b}_v =
(b_{v,1}, \ldots, b_{v,K})$ over $K$ candidate states (including an
``unknown'' state). The uncertainty of $v$ is quantified by Shannon entropy:
\[
  H(v) = -\sum_{k=1}^{K} b_{v,k} \log_2 b_{v,k}.
\]
An entity is \emph{resolved} when $H(v) \leq \delta$ for a resolution
threshold $\delta \geq 0$. The \emph{importance} of unresolved entity $v$ is:
\[
  I(v) = w(v) \cdot H(v).
\]
\end{definition}

\begin{definition}[Coverage Function]
\label{def:coverage}
For a set of resolved entities $S \subseteq V$, the weighted coverage is:
\[
  C(S) = \frac{\sum_{v \in S} w(v)}{\sum_{v \in V} w(v)}.
\]
\end{definition}

\begin{definition}[Frontier]
\label{def:frontier}
The \emph{frontier} at time $t$ is the set of entities whose prerequisites are
all resolved but which are themselves unresolved:
\[
  \mathcal{F}^{(t)} = \bigl\{v \in V \setminus S^{(t)} :
    \forall u \text{ s.t.\ } (u, v) \in E_P, \; u \in S^{(t)} \bigr\},
\]
where $S^{(t)}$ is the set of resolved entities at time $t$. Only frontier
entities are eligible for querying.
\end{definition}

% ----------------------------------------------------------------------------
\subsection{Neighbor-conditioned stability}
\label{sec:theory_ncs}

We now introduce the central theoretical concept that distinguishes
DualGraph from flat or ad-hoc task representations. The key observation
is that real information-gathering tasks exhibit \emph{local dependency
structure}: the importance of querying a particular entity depends not on
the global state of all entities but only on the states of a small set
of logically or semantically related entities---its graph neighbors.

\begin{definition}[Neighborhood and Neighbor-Conditioned Stability]
\label{def:ncs}
For a node $v \in V$, define its \emph{combined neighborhood} as the union
of its prerequisite parents and association neighbors:
\[
  \mathcal{N}(v) = \bigl\{u \in V : (u, v) \in E_P\bigr\}
    \cup \bigl\{u \in V : \{u, v\} \in E_A\bigr\}.
\]
A scoring function $\sigma: V \times 2^V \to \mathbb{R}$ satisfies
\emph{neighbor-conditioned stability} (NCS) if for all $v \in V$ and
all time steps $t, t'$:
\[
  \bigl[\forall u \in \mathcal{N}(v): \mathbf{b}_u^{(t')} =
    \mathbf{b}_u^{(t)}\bigr]
  \implies
  \sigma(v, S^{(t')}) = \sigma(v, S^{(t)}).
\]
That is, if the belief states of all neighbors of $v$ are unchanged between
turns $t$ and $t'$, then the score of $v$ is also unchanged, regardless of
changes to distant nodes.
\end{definition}

\begin{proposition}[DualGraph Scoring Satisfies NCS]
\label{prop:ncs_holds}
The DualGraph scoring function
\[
  \mathrm{Score}(v) = \alpha \cdot \tilde{I}(v)
    + \beta \cdot T(v) + \gamma \cdot C(v)
\]
satisfies neighbor-conditioned stability under the following conditions:
(i) the importance term $\tilde{I}(v)$ depends on $v$'s belief and the
beliefs of its prerequisite parents (which determine frontier membership);
(ii) the centrality term $T(v)$ is precomputed from the static graph
topology and does not change during dialogue; (iii) the community continuity
term $C(v)$ depends only on the previously queried node $v_{\mathrm{prev}}$,
which is a global quantity but changes at most once per turn and affects all
nodes equally.
\end{proposition}

\begin{proof}
Consider the three terms independently. The normalized importance
$\tilde{I}(v) = I(v) / \max_{u \in \mathcal{F}} I(u)$ depends on $v$'s own
entropy $H(v)$ (which changes only when $v$ is directly queried) and on
frontier membership (which changes only when a prerequisite parent of $v$
transitions from unresolved to resolved). Both conditions involve
$\mathcal{N}(v)$. The PageRank centrality $T(v)$ is a function of the static
graph topology $G_A$ and is invariant across turns. The community continuity
$C(v) = \mathbf{1}[\mathrm{comm}(v) = \mathrm{comm}(v_{\mathrm{prev}})]$
depends on $v$'s community assignment (static) and on $v_{\mathrm{prev}}$,
which is a global parameter. Conditioning on $v_{\mathrm{prev}}$ being fixed
within a turn, $C(v)$ is also locally determined. Therefore,
$\mathrm{Score}(v)$ changes only when (a) $v$'s own belief changes,
(b) a prerequisite parent's belief changes (affecting frontier status), or
(c) an association neighbor's belief changes (affecting the normalization
denominator if the neighbor is also in the frontier). All three conditions
involve nodes in $\mathcal{N}(v)$.
\end{proof}

\begin{remark}[Connection to Dynamic Programming]
\label{rem:dp}
The NCS property is structurally analogous to the \emph{principle of
optimality} in dynamic programming~\cite{bellman1957dynamic}. In the Bellman
equation, the optimal value of a state depends only on the values of its
successor states, not on the entire state space. This locality enables
efficient value iteration: only states whose successors have changed need
recomputation. In DualGraph, the optimal next-query decision depends only
on the states of graph-local neighbors, enabling efficient priority updates
that propagate only through the graph's edge structure rather than requiring
global recomputation at every turn.

This connection runs deeper than mere analogy. Consider the dual graphs
as defining a state-transition structure over entity configurations. At each
turn, the agent transitions from one configuration (set of resolved entities)
to another by resolving one additional entity. The ``value'' of each
configuration is the expected future cost (remaining turns) to achieve full
coverage. The NCS property ensures that this value function decomposes
locally over the graph, enabling a greedy policy that approximates the
optimal policy within the bounds established by submodularity
(Theorem~\ref{thm:convergence}).

More precisely, let $\mathcal{S} = 2^V$ be the space of entity
configurations (resolved subsets). For a given configuration $S$, the
Bellman equation for the minimum expected queries to coverage
$\geq 1 - \epsilon$ is:
\[
  Q^*(S) = \begin{cases}
    0 & \text{if } C(S) \geq 1 - \epsilon, \\
    1 + \min_{v \in \mathcal{F}(S)} \bigl[p_v \cdot Q^*(S \cup \{v\})
      + (1 - p_v) \cdot Q^*(S)\bigr] & \text{otherwise.}
  \end{cases}
\]
The NCS property implies that for any two configurations $S, S'$ that differ
only in entities \emph{outside} $\mathcal{N}(v)$, the marginal value of
querying $v$ is identical:
\[
  Q^*(S) - Q^*(S \cup \{v\}) = Q^*(S') - Q^*(S' \cup \{v\}),
\]
provided $\mathcal{F}(S) \cap \mathcal{N}(v) = \mathcal{F}(S') \cap
\mathcal{N}(v)$. This is precisely the condition under which the greedy
policy (always query the highest-score frontier entity) achieves near-optimal
performance, because the submodularity of $C$ ensures that greedy marginal
gains are good proxies for global optimality.
\end{remark}

The practical consequence of NCS is threefold. First, \emph{computational
efficiency}: at each turn, only the $O(\Delta)$ neighbors of the most
recently updated entity (where $\Delta = \max_v |\mathcal{N}(v)|$) require
score recomputation, rather than all $|V|$ entities. For typical task graphs
where $\Delta \ll |V|$, this is a substantial reduction. Second,
\emph{consistency}: because scores are updated only in response to actual
neighbor state changes, the system cannot enter oscillatory states where an
entity's priority flickers due to irrelevant distant changes. Third,
\emph{interpretability}: the locality of updates means that any change in
agent behavior (e.g., switching from one topic to another) can be traced to a
specific neighboring entity's state change, providing an auditable explanation.

\begin{proposition}[Local Update Sufficiency]
\label{prop:local_update}
Under the DualGraph scoring function (Eq.~\ref{eq:score}), the optimal
next-entity selection at turn $t+1$ depends only on the set of nodes whose
combined neighborhoods intersect with the set of entities updated at turn $t$.
Formally, let $\Delta_t \subseteq V$ be the set of entities whose belief
distributions changed at turn $t$. Define the \emph{update frontier}:
\[
  \mathcal{U}_t = \bigl\{v \in V : \mathcal{N}(v) \cap \Delta_t
    \neq \emptyset\bigr\} \cup \Delta_t.
\]
Then $\mathrm{Score}^{(t+1)}(v) = \mathrm{Score}^{(t)}(v)$ for all
$v \notin \mathcal{U}_t$.
\end{proposition}

\begin{proof}
Follows directly from Definition~\ref{def:ncs} and
Proposition~\ref{prop:ncs_holds}. For $v \notin \mathcal{U}_t$, no node in
$\mathcal{N}(v)$ has changed belief state (since
$\mathcal{N}(v) \cap \Delta_t = \emptyset$ and $v \notin \Delta_t$),
so $\mathrm{Score}(v)$ is unchanged by NCS.
\end{proof}

% ----------------------------------------------------------------------------
\subsection{Submodularity and convergence guarantees}
\label{sec:theory_convergence}

\begin{proposition}[Submodularity of Weighted Coverage]
\label{prop:submodular}
The weighted coverage function $C: 2^V \to [0,1]$ is monotone submodular.
\end{proposition}

\begin{proof}
Monotonicity is immediate: for $S \subseteq T$, $C(S) \leq C(T)$ since
adding resolved entities can only increase the weighted sum in the numerator.
For submodularity, consider $A \subseteq B \subseteq V$ and $v \notin B$. Then
\[
  C(A \cup \{v\}) - C(A) = \frac{w(v)}{W} = C(B \cup \{v\}) - C(B),
\]
where $W = \sum_{u \in V} w(u)$. The marginal gain is constant (does not
decrease), so $C$ is in fact modular. However, in the
prerequisite-constrained setting, the coverage function is defined only over
feasible sets (those respecting topological order in $G_P$). The restriction
of a modular function to a matroid constraint yields a submodular optimization
problem over a matroid~\cite{calinescu2011, vondrak2008}, for which the greedy
algorithm achieves a $(1 - 1/e)$-approximation.
\end{proof}

\begin{theorem}[Convergence Bound for DualGraph Greedy Policy]
\label{thm:convergence}
Let $G_P = (V, E_P)$ be a DAG with longest path length $L$ and let
$\Delta_{\max} = \max_v |\mathcal{N}(v)|$ be the maximum combined
neighborhood size. Assume each query resolves entity $v$ with probability
$p_{\min} > 0$. Then DualGraph's greedy policy achieves
$C(S) \geq 1 - \epsilon$ in at most
\[
  T \leq \frac{|V|}{p_{\min}} \left(1 +
    \frac{\ln(1/\epsilon)}{\ln\bigl(1/(1 - p_{\min})\bigr)}\right)
\]
queries in expectation, respecting all prerequisite constraints. Furthermore,
the per-turn computational cost of the NCS-based score update is
$O(\Delta_{\max})$, yielding a total computational cost of
$O(T \cdot \Delta_{\max})$ for the complete dialogue.
\end{theorem}

\begin{proof}
\placeholder{%
  \textbf{Proof sketch (full proof in Supplementary~\ref{app:proof}):}
  \begin{enumerate}[noitemsep]
    \item The prerequisite graph $G_P$ is a DAG; topological layers
          $L_0, L_1, \ldots, L_d$ (where $d \leq L$) partition $V$ such
          that all prerequisites of entities in $L_i$ lie in $L_0 \cup
          \cdots \cup L_{i-1}$.
    \item At any time $t$, the frontier $\mathcal{F}^{(t)}$ contains only
          entities whose layer predecessors are all resolved. The greedy
          policy queries the highest-importance frontier entity.
    \item Each query resolves the target entity with probability
          $\geq p_{\min}$; after $k$ attempts, resolution probability is
          $1 - (1 - p_{\min})^k \geq 1 - e^{-k p_{\min}}$.
    \item Setting $k^* = \lceil \ln(|V|/\epsilon) / p_{\min} \rceil$ ensures
          that each entity is resolved with probability $\geq 1 - \epsilon / |V|$.
    \item By a union bound over $|V|$ entities and summing over at most
          $L + 1$ layers, the total expected queries is bounded by
          $|V| \cdot k^*$, yielding the stated bound.
    \item The $(1 - 1/e)$ submodularity guarantee applies to the
          importance-weighted ordering within each topological layer.
    \item The NCS property ensures that only $O(\Delta_{\max})$ scores are
          recomputed per turn (Proposition~\ref{prop:local_update}).
  \end{enumerate}
}
\end{proof}

\begin{corollary}[When Dual-Graph Outperforms Single-Graph and Flat Methods]
\label{cor:when}
The performance advantage of DualGraph over alternative control structures
can be characterized as follows.

\noindent\textbf{(a) Over prerequisite-only control.} Let $\phi \in [0, 1]$
denote the \emph{oracle cooperation probability}, modeled as a decreasing
function of topic incoherence: $\phi = g(\mathrm{coherence})$ where coherence
is measured by community continuity in $G_A$. Without the association graph,
topic transitions are unconstrained, yielding lower $\phi$ and thus lower
effective $p_v$. The convergence bound of Theorem~\ref{thm:convergence}
depends on $p_{\min}$; reducing $p_{\min}$ by a factor of $\phi < 1$
increases the query bound by a factor of $1/\phi$.

\noindent\textbf{(b) Over association-only control.} Without prerequisite
constraints, the agent may query entities whose prerequisites are unresolved,
yielding uninformative responses ($p_v \approx 0$ for such queries). The
expected number of wasted queries is proportional to the number of
prerequisite-violating selections, which scales with the depth $L$ of the
prerequisite DAG.

\noindent\textbf{(c) Over flat checklists.} A checklist provides no ordering
and no coherence structure. It is equivalent to a DualGraph with
$E_P = E_A = \emptyset$, yielding $\Delta_{\max} = 0$ and no locality to
exploit. Both prerequisite violations and incoherence penalties apply,
compounding into a query bound that is $O(L / \phi)$ times worse than
DualGraph in the worst case.

\placeholder{%
  Formalize the oracle cooperation model $\phi = g(\cdot)$.
  Provide empirical calibration of $\phi$ from simulation data.
  Plot theoretical bound vs.\ empirical convergence for each condition.
}
\end{corollary}

% ----------------------------------------------------------------------------
\subsection{Complexity analysis}
\label{sec:theory_complexity}

The per-turn computational overhead of DualGraph beyond base LLM inference
consists of four components. Score recomputation under NCS costs
$O(\Delta_{\max})$ per turn, where $\Delta_{\max}$ is the maximum combined
neighborhood size. Community detection via the Leiden
algorithm~\cite{traag2019leiden} costs $O(|E_A|)$ and is executed once at
initialization and only re-run if the graph topology changes (which it does
not in the current static-graph formulation). Memory retrieval using
approximate nearest-neighbor indexing (HNSW) costs $O(d \log |M|)$ per
query, where $|M|$ is the memory store size and $d$ is the embedding
dimension. Entity extraction from the user response requires one additional
LLM call (or a lightweight NER model), costing $O(1)$ relative to the main
dialogue LLM call. For typical task graphs with $|V| \leq 100$ entities and
$\Delta_{\max} \leq 15$, the total per-turn overhead is negligible compared
to LLM inference latency.

\placeholder{%
  \textbf{Table: Per-turn latency breakdown (milliseconds).}\\
  \begin{tabular}{lrr}
    Component & DualGraph & Global recompute (no NCS) \\
    Score update & [X] ms & [Y] ms \\
    Memory retrieval & [X] ms & [X] ms \\
    Entity extraction & [X] ms & [X] ms \\
    LLM inference & [X] ms & [X] ms \\
    Total & [X] ms & [X] ms \\
  \end{tabular}
}



% ----------------------------------------------------------------------------
\subsection{Evolving inference DAGs: online graph extension under NCS}
\label{sec:theory_evolving}

We now formalize the extension from static task graphs to dynamically evolving
inference DAGs that grow during dialogue as new information entities are
discovered.

\begin{definition}[Evolving Inference DAG]
\label{def:evolving}
An \emph{evolving inference DAG} is a time-indexed sequence of structured
information-gathering problems $\mathcal{T}^{(t)} = (V^{(t)}, E_P^{(t)},
E_A^{(t)}, w^{(t)}, \mathcal{O})$ where:
\begin{itemize}[noitemsep]
  \item $V^{(0)} \subseteq V^{(1)} \subseteq \cdots$ (the node set can
        only grow);
  \item At each turn $t$, at most one new node $v_{\mathrm{new}}$ may be
        added: $V^{(t+1)} = V^{(t)} \cup \{v_{\mathrm{new}}\}$;
  \item New prerequisite edges $(u, v_{\mathrm{new}})$ and
        $(v_{\mathrm{new}}, u)$ are added subject to the constraint that
        $G_P^{(t+1)}$ remains a DAG;
  \item New association edges $\{u, v_{\mathrm{new}}\}$ are added based on
        embedding similarity exceeding threshold $\theta$;
  \item $w^{(t+1)}(v_{\mathrm{new}})$ is assigned by the importance
        estimation module (LLM-based or rule-based).
\end{itemize}
The coverage function generalizes to the time-varying node set:
$C^{(t)}(S) = \sum_{v \in S \cap V^{(t)}} w^{(t)}(v) \big/
\sum_{v \in V^{(t)}} w^{(t)}(v)$.
\end{definition}

\begin{proposition}[NCS Containment for Online Node Attachment]
\label{prop:evolving_ncs}
When a new node $v_{\mathrm{new}}$ is attached to the graph at turn $t+1$,
the set of existing nodes whose scores may change is contained within:
\[
  \mathcal{U}_{t+1}^{\mathrm{attach}} = \mathcal{N}(v_{\mathrm{new}})
    \cup \bigl\{u \in V^{(t)} : v_{\mathrm{new}} \in \mathcal{N}(u)
    \text{ in } G^{(t+1)}\bigr\}.
\]
That is, only the direct neighbors of the new node in the updated graph
require score recomputation. All nodes $v$ with
$\mathcal{N}^{(t+1)}(v) \cap \{v_{\mathrm{new}}\} = \emptyset$ retain
their scores from turn $t$.
\end{proposition}

\begin{proof}
The NCS property (Definition~\ref{def:ncs}) states that
$\mathrm{Score}(v)$ is unchanged when no neighbor's belief state has
changed. Adding $v_{\mathrm{new}}$ introduces a new neighbor only for
nodes directly connected to $v_{\mathrm{new}}$ by a prerequisite or
association edge in $G^{(t+1)}$. For all other nodes, the neighborhood
$\mathcal{N}^{(t+1)}(v) = \mathcal{N}^{(t)}(v)$ and no belief states
within that neighborhood have changed. The PageRank centrality $T(v)$ is
recomputed only if the topology change affects $v$'s subgraph; for nodes
far from $v_{\mathrm{new}}$, the change in PageRank is negligible and can
be bounded by the mixing-time analysis of personalized
PageRank~\cite{page1999pagerank}. Community assignments may change if
$v_{\mathrm{new}}$ bridges two communities; we handle this by running
incremental community detection only on the affected community and its
neighbors.
\end{proof}

\begin{theorem}[Convergence of Evolving DualGraph]
\label{thm:evolving_convergence}
Let $V^{(\infty)} = \bigcup_{t \geq 0} V^{(t)}$ be the final node set
(which is finite since the dialogue has finite length and each turn adds at
most one node). Assume $|V^{(\infty)}| \leq N_{\max}$. Then the convergence
bound of Theorem~\ref{thm:convergence} applies with $|V|$ replaced by
$N_{\max}$, and the per-turn computational cost remains
$O(\Delta_{\max}^{(\infty)})$ where $\Delta_{\max}^{(\infty)}$ is the
maximum neighborhood size in the final graph.
\end{theorem}

\begin{proof}[Proof sketch]
The evolving DAG only adds nodes and edges, never removes them. At any
turn $t$, the subgraph over $V^{(t)}$ is a valid structured
information-gathering problem satisfying all requirements of
Definition~\ref{def:sig} (since DAG-ness is maintained by construction).
Theorem~\ref{thm:convergence} applies to this subgraph. Because
$V^{(t)} \subseteq V^{(\infty)}$ and the bound is monotone in $|V|$,
the worst-case bound uses $N_{\max}$. NCS containment
(Proposition~\ref{prop:evolving_ncs}) ensures that per-turn cost does not
exceed $O(\Delta_{\max}^{(\infty)})$ regardless of when nodes are added.
\end{proof}

\begin{remark}[DAG Maintenance Under Online Growth]
\label{rem:dag_maintenance}
A critical implementation constraint is that prerequisite edges involving
$v_{\mathrm{new}}$ must not introduce cycles into $G_P$. When the online
graph-extension module proposes a prerequisite edge
$(v_{\mathrm{new}}, u)$ for some existing node $u$, the system verifies
that $u$ is not an ancestor of $v_{\mathrm{new}}$ in the current $G_P$.
This check costs $O(|V^{(t)}|)$ in the worst case (a single DFS from $u$)
but is fast in practice because the prerequisite graph is sparse. If the
proposed edge would create a cycle, it is rejected and the LLM-based
dependency assessor is re-queried with the constraint made explicit.
\end{remark}


\subsubsection{Evolving inference DAG implementation}
\label{sec:methods_evolving}

The evolving-DAG mechanism operates through a three-stage pipeline triggered
whenever the entity extraction module detects information that does not map
to any existing node in $V^{(t)}$ with embedding cosine similarity above a
matching threshold $\theta_{\mathrm{match}}$ (default 0.85).

\paragraph{Stage 1: Entity emergence detection.}
After each user response, the extraction module produces a set of candidate
entity--value pairs. Each candidate entity description is embedded and
compared (cosine similarity) against all existing node descriptions in
$V^{(t)}$. If the maximum similarity falls below $\theta_{\mathrm{match}}$,
the entity is flagged as potentially emergent. To reduce false positives,
the candidate is then assessed by a lightweight LLM call: ``Does the
following information represent a new clinical/task-relevant entity not
covered by any of the following existing entities? [list of top-5 most
similar existing entities]''. Only candidates confirmed by both the
embedding filter and the LLM check proceed to Stage 2.

\paragraph{Stage 2: Graph attachment.}
For a confirmed emergent entity $v_{\mathrm{new}}$, the system constructs
its local graph connections. Prerequisite edges are determined by prompting
the LLM: for each existing node $u$ in the top-$k$ most similar nodes
(default $k = 10$), the system asks ``Must $u$ be resolved before
$v_{\mathrm{new}}$ can be meaningfully queried?'' and vice versa. Edges
with positive responses are added subject to DAG verification (a DFS-based
cycle check). Association edges are added for all existing nodes with
embedding similarity above $\theta$ (same threshold as initial graph
construction). The importance weight $w(v_{\mathrm{new}})$ is estimated
by the LLM on a 1--5 scale, consistent with the initial graph construction
pipeline.

\paragraph{Stage 3: NCS-contained propagation.}
Once $v_{\mathrm{new}}$ and its edges are committed to the graph, the
NCS dirty-flag mechanism is invoked: dirty flags are set for all nodes in
$\mathcal{N}(v_{\mathrm{new}})$ (the combined neighborhood in the updated
graph). Community detection is run incrementally: only the community
containing $v_{\mathrm{new}}$'s closest association neighbor is
re-evaluated, using the Leiden algorithm seeded with the current partition.
PageRank is updated via a rank-one perturbation approximation rather than
full recomputation. All other graph regions remain untouched.

The entire pipeline executes within a single dialogue turn. The
computational overhead beyond the standard DualGraph per-turn cost is one
additional LLM call (emergence verification), $k$ LLM calls (prerequisite
assessment), and $O(|V^{(t)}|)$ embedding comparisons for association
edges. For typical emergence rates (fewer than [X] new entities per
dialogue) and graph sizes ($|V| \leq 100$), this adds \placeholder{[Y] ms}
per emergence event.

\placeholder{%
  \textbf{Failure modes and safeguards:}
  \begin{itemize}[noitemsep]
    \item \textbf{False-positive emergence}: a previously existing entity
          is misidentified as new, creating a duplicate node. Safeguard:
          after graph attachment, a post-hoc deduplication check compares
          the new node's resolved value against all existing nodes; if a
          match is found, the nodes are merged.
    \item \textbf{False-negative emergence}: a genuinely new entity is
          matched to an existing node. Safeguard: if the existing node's
          belief distribution becomes bimodal after the new information
          is integrated (entropy increases above $\delta$), the system
          re-evaluates whether a split into two nodes is warranted.
    \item \textbf{Incorrect prerequisite assignment}: a spurious
          prerequisite edge blocks the new node or an existing node.
          Safeguard: if the new node remains unresolved after $h_{\max}$
          attempts, its prerequisite edges are re-evaluated.
    \item \textbf{Runaway graph growth}: a pathological user provides
          unlimited novel information, causing the graph to grow without
          bound. Safeguard: a maximum graph size $N_{\max}$ (default 150)
          caps the number of nodes; beyond this, new entities are stored
          in memory but not added to the graph.
  \end{itemize}
}
% ============================================================================



% 4. DISCUSSION
% ============================================================================
\section{Discussion}
\label{sec:discussion}

% --- Paragraph 1: Key finding and conceptual contribution ---
The central finding of this work is that structured information gathering
is a distinct capability from memory and from general planning, and current
LLM agent architectures lack the appropriate primitive for it. Memory
enables retention; planning enables action sequencing; but neither ensures
systematic, dependency-aware coverage of a defined information space.
DualGraph cleanly separates three architectural primitives: memory
(retention of acquired information), task control (determining what to
ask next, subject to structural constraints), and interaction management
(formulating questions coherently). The dual-graph formulation, grounded
in the neighbor-conditioned stability principle, provides both formal
guarantees and practical effectiveness across diverse domains.

% --- Paragraph 2: Why graphs, and the NCS insight ---
The theoretical contribution of neighbor-conditioned stability provides a
deeper answer to the question ``why graphs?''\ that goes beyond the
operational argument that graphs conveniently encode prerequisites. Graphs
are the \emph{correct} abstraction for information-gathering tasks because
such tasks inherently possess \emph{local dependency structure}: whether a
particular entity needs to be queried depends not on the global conversation
state but on the states of its logically and semantically adjacent entities.
This locality is precisely what graph edges encode and what flat
representations---checklists, priority queues, linear plans---discard. The
NCS principle makes this locality explicit and actionable: it guarantees
that the agent's behavior is stable under distant state changes, preventing
the oscillatory and inconsistent behavior observed in globally-reasoning
agents that attempt to re-evaluate all priorities from scratch at every turn.

The connection to dynamic programming further clarifies the framework's
theoretical standing. Just as Bellman's principle of optimality decomposes
a global optimization over an exponentially large state space into local
value updates that propagate along state-transition edges, DualGraph
decomposes the global information-gathering optimization into local
priority updates that propagate along prerequisite and association edges.
The greedy policy inherits near-optimality guarantees from the submodularity
of coverage, and NCS ensures that these guarantees are achieved with local
(rather than global) computation per turn.

% --- Paragraph 3: Why dual graphs, not single graph ---
The ablation analysis reveals that prerequisite and association edges serve
complementary and non-interchangeable functions: prerequisites ensure the
\emph{logical validity} of questions (asking about nephropathy before
confirming renal function is meaningless), while associations ensure
\emph{conversational coherence} (jumping from medication history to family
social dynamics is jarring and reduces oracle cooperation). Merging them
into a single graph degrades both properties because the merged edges
conflate ``must come before'' with ``is related to,'' losing the directional
semantics of prerequisites and the symmetric semantics of association.
This separation is analogous to the distinction between control flow and
data flow in programming language design: both are essential, but collapsing
them produces a less expressive and less analyzable structure.

% --- Paragraph 4: Generality and graph construction ---
The semi-automatic graph construction pipeline demonstrates that the
framework's requirements are not prohibitively domain-specific. Task graphs
achieving [X]\% F1 against expert-designed ground truth can be generated
automatically from textual task specifications, with expert review requiring
only [Y] minutes per domain. Downstream performance using auto-generated
graphs is within [Z]\% of expert-designed graphs, suggesting that the
framework is robust to moderate graph imperfections---a consequence of the
NCS principle, which localizes the effect of any single erroneous edge to
the immediate neighborhood. This robustness is confirmed by the
perturbation experiments (Section~\ref{sec:methods_robustness}), which
show graceful degradation rather than catastrophic failure under edge
deletion, addition, and reversal.

% --- Paragraph 5: The memory quality side-effect ---
An unexpected finding is that DualGraph improves not only task completion
but also memory quality, as evidenced by the LoCoMo benchmark results
(Table~\ref{tab:locomo}). Because the graph controller feeds
entity-tagged, confidence-scored, structurally organized information into
memory rather than raw dialogue text, retrieval precision improves as a
side-effect. The NCS mechanism further contributes by ensuring that stored
entity states are consistent and stable: entities are updated only when
their neighborhoods change, reducing contradictory overwrites that degrade
multi-hop and temporal query performance. This suggests that task control
and memory quality are synergistic rather than independent---a finding
with implications for memory system design more broadly.

% --- Paragraph 6: Limitations (revised: evolving DAG partially addresses
%     the old limitation about pre-defined entity sets) ---
Several limitations should be noted. Although the evolving-DAG extension
substantially relaxes the closed-world assumption of static task graphs,
it still requires that the initial graph cover the \emph{majority} of
expected entities; the online-discovery mechanism is designed for
occasional emergent entities, not for tasks where the information space
is entirely unknown at the outset. Open-ended tasks such as creative
brainstorming, psychotherapy, or exploratory research interviews, where
the relevant entities emerge dynamically and constitute the majority of
the information space, fall outside the current scope and would require a
fundamentally different graph-seeding mechanism. The oracle cooperation
model assumes a generally cooperative interlocutor; adversarial or
deliberately deceptive settings would require additional mechanisms. The
prospective pilot is small-scale and single-center, providing feasibility
evidence rather than definitive clinical validation. Computationally,
while the NCS-based update is efficient for typical graph sizes
($|V| \leq 100$), very large entity sets ($|V| > 1000$) or high-frequency
node attachment (multiple new entities per turn) would benefit from
hierarchical graph decomposition and batched NCS updates. We evaluate with
one base LLM family (Qwen-2.5-32B); generalization to other architectures
should be verified. The evolving-DAG mechanism relies on the extraction
module's ability to detect that user-provided information does not map to
any existing node---a classification problem that may produce false
positives (spurious node creation) or false negatives (missed emergent
entities); we report both rates in the evaluation but do not claim
optimal detection. Finally, the theoretical analysis assumes independent
entity resolution probabilities, whereas in practice, resolving one entity
may change the probability of resolving others through trust-building,
contextual priming, or information leakage.

% --- Paragraph 7: Broader implications ---
As LLM agents are deployed in high-stakes information-gathering
settings---clinical intake, legal discovery, financial due diligence,
regulatory compliance---the gap between ``can remember'' and ``will
systematically complete'' becomes safety-critical. An agent that remembers
a patient's name but fails to ask about drug allergies creates a false
sense of competence. DualGraph provides a principled, auditable mechanism
for task control: the graph structure is inspectable, the NCS updates are
traceable, and the coverage metric is quantifiable. The framework
complements rather than competes with advances in memory management and
reasoning capabilities.

% --- Paragraph 8: Future work (revised: evolving DAG is now achieved;
%     new future directions) ---
Several directions for future work emerge from this framework. The most
immediate is \emph{hierarchical evolving DAGs} for tasks with nested
sub-goals: a top-level DAG over goal categories, each containing a
sub-DAG over specific entities, with NCS operating at both levels.
A second direction is \emph{multi-agent coordination}: multiple
DualGraph-controlled agents sharing or partitioning a single task graph,
with NCS providing the locality guarantee that one agent's updates do not
invalidate another agent's traversal plan---a property that maps naturally
to distributed dynamic programming.
A third direction is \emph{adversarial robustness}: extending the oracle
cooperation model to account for deliberately misleading responses,
integrating contradiction detection into the NCS update rule so that
inconsistent information triggers neighborhood re-evaluation.
A fourth is integration with \emph{multimodal inputs}---clinical images,
sensor data, uploaded documents---that resolve entities through
non-conversational channels; the evolving-DAG mechanism already provides
the infrastructure for attaching new nodes discovered through non-textual
modalities.
Finally, larger prospective studies with clinical outcome endpoints---blood
pressure reduction, medication adherence, patient satisfaction, and
time-to-treatment---would establish the clinical impact that justifies
deployment at scale.

% ============================================================================
% 5. METHODS
% ============================================================================
\section{Methods}
\label{sec:methods}

\subsection{DualGraph architecture}
\label{sec:methods_architecture}

\subsubsection{Dual-graph representation}

The entity prerequisite graph $G_P = (V, E_P)$ is a directed acyclic graph
encoding logical dependency among required information entities. The
association graph $G_A = (V, E_A, w_A)$ is an undirected weighted graph
encoding semantic relatedness among the same entities for coherent topic
transitions. Nodes in $V$ correspond to atomic information units required for
the target task; each node maintains the belief distribution, entropy, and
metadata described in Definitions~\ref{def:entropy} and~\ref{def:ncs}. Edges
in $E_P$ represent hard prerequisite constraints (entity $u$ must be resolved
before entity $v$ can be meaningfully queried). Edges in $E_A$ represent
semantic association, with weights $w_A(u, v)$ proportional to the cosine
similarity of entity description embeddings.

The two graphs share the same node set $V$ but encode complementary
information: $G_P$ encodes the ``must come before'' relation (directed,
acyclic), while $G_A$ encodes the ``is related to'' relation (undirected,
potentially cyclic). This separation is essential: merging them would lose
the directional semantics of prerequisites and the community structure of
associations.

\subsubsection{Neighbor-conditioned stability implementation}

At each dialogue turn, after the user response is processed and entity
beliefs are updated, the NCS mechanism determines which node scores require
recomputation. Let $\Delta_t \subseteq V$ be the set of entities whose belief
distributions changed at turn $t$ (typically $|\Delta_t| = 1$, the entity
that was directly queried, though it may include additional entities if the
user's response provides information about multiple entities). The update
frontier $\mathcal{U}_t = \{v \in V : \mathcal{N}(v) \cap \Delta_t \neq
\emptyset\} \cup \Delta_t$ is computed, and scores are recomputed only for
$v \in \mathcal{U}_t$. All other scores are guaranteed unchanged by NCS
(Proposition~\ref{prop:local_update}).

In practice, this is implemented as a \texttt{dirty flag} mechanism: each
node maintains a boolean flag that is set to \texttt{true} when any of its
neighbors' beliefs change. Only flagged nodes are re-scored; flags are
cleared after re-scoring. This reduces per-turn score computation from
$O(|V|)$ to $O(|\mathcal{U}_t|) = O(\Delta_{\max})$.

\subsubsection{Node scoring and selection}

At each dialogue turn, the controller selects the next entity to query from
the frontier $\mathcal{F}^{(t)}$ (Definition~\ref{def:frontier}). For each
frontier entity $v$, the composite score is:
\begin{equation}
  \mathrm{Score}(v) = \alpha \cdot \tilde{I}(v) + \beta \cdot T(v)
    + \gamma \cdot C(v),
  \label{eq:score}
\end{equation}
where $\tilde{I}(v) = I(v) / \max_{u \in \mathcal{F}} I(u)$ is normalized
importance over the current frontier, $T(v)$ is the precomputed PageRank
centrality of $v$ in $G_A$ (capturing structural importance: nodes central in
the association graph are topic hubs that connect multiple information
clusters), and $C(v) = \mathbf{1}[\mathrm{comm}(v) =
\mathrm{comm}(v_{\mathrm{prev}})]$ is a binary community continuity score
that favors entities in the same semantic community as the most recently
queried entity, promoting topically coherent conversation flow.

Community structure is detected by applying the Leiden
algorithm~\cite{traag2019leiden} to $G_A$ at initialization. The resolution
parameter is set to [value] to yield communities of [typical size] entities,
corresponding to natural topic clusters.

\placeholder{%
  \textbf{Specify parameter values:}
  $\alpha = $ [value], $\beta = $ [value], $\gamma = $ [value].\\
  \textbf{Hit counters:} Each node maintains a counter $h_v$ tracking how
  many times it has been queried. If $h_v \geq h_{\max}$ (default [value]),
  the node is deprioritized by multiplying $\mathrm{Score}(v)$ by a decay
  factor $\rho^{h_v - h_{\max}}$ (default $\rho = $ [value]). This prevents
  the agent from repeatedly querying an entity that the user cannot or will
  not answer, while still allowing a limited number of re-attempts for
  partially informative responses.
}

\subsubsection{Information extraction and confidence gating}

After each user response, an extraction module parses the response to identify
entity values. Each extracted value is assigned a confidence score
$\rho_v \in [0,1]$ based on directness (explicitly stated vs.\ inferred),
consistency (agreement with previously stored values), and specificity (precise
values vs.\ vague ranges). Values with $\rho_v < \rho_{\min}$ (default 0.6)
are flagged as unreliable and trigger targeted clarification at the next
appropriate opportunity (respecting NCS: the low-confidence extraction sets
the entity's dirty flag, ensuring it is reconsidered) rather than being
committed to memory.

Belief distributions are updated via a Bayesian update rule:
\begin{equation}
  b_{v,k}^{(t+1)} = \frac{b_{v,k}^{(t)} \cdot \ell(o_v \mid s_k)}
    {\sum_{j=1}^{K} b_{v,j}^{(t)} \cdot \ell(o_v \mid s_j)},
  \label{eq:belief_update}
\end{equation}
where $\ell(o_v \mid s_k)$ is the likelihood of the observed response $o_v$
given that entity $v$ is in state $s_k$.

\placeholder{%
  \textbf{Specify likelihood model:} How is $\ell(o_v \mid s_k)$
  computed in practice? Options:
  \begin{itemize}[noitemsep]
    \item LLM-based: prompt the LLM to assess the probability that the
          user's response indicates state $s_k$. Calibrate with
          temperature scaling.
    \item Rule-based: for structured entities (blood pressure, medication
          names), use pattern matching with predefined state categories.
    \item Hybrid: rule-based for structured entities, LLM-based for
          free-text entities.
  \end{itemize}
}

\subsubsection{Long-term memory integration}

Confirmed entities (those with $\rho_v \geq \rho_{\min}$ and
$H(v) \leq \delta$) are stored as structured key-value pairs in a persistent
memory store indexed by entity embeddings. At each turn, the top-$m$ most
relevant memory items (by cosine similarity to the current query context and
the active entity's neighborhood in $G_A$) are retrieved and injected into
the prompt context, providing the LLM with relevant previously acquired
information for formulating contextually appropriate follow-up questions.

\placeholder{%
  \textbf{Specify memory architecture details:}
  \begin{itemize}[noitemsep]
    \item Storage format: JSON-structured entity records with fields:
          entity\_id, entity\_name, value, confidence, timestamp, source\_turn,
          evidence\_snippet, belief\_distribution.
    \item Embedding model: BAAI/bge-large-[language]-v1.5.
    \item Retrieval: HNSW approximate nearest-neighbor index; top-$m$
          (default $m = 5$).
    \item Conflict resolution: when a new extraction conflicts with a stored
          value, both are retained with timestamps; the higher-confidence
          value is used for downstream reasoning, but the conflict is flagged
          for potential clarification. This design leverages NCS: the
          conflicting entity's dirty flag is set, ensuring its score
          increases (due to increased entropy) and it is re-queried for
          clarification.
    \item Capacity management: [describe eviction or summarization policy if
          memory grows too large].
  \end{itemize}
}

\subsection{Semi-automatic graph construction}
\label{sec:graph_construction}

The knowledge embedded in task specifications—clinical guidelines,
regulatory forms, troubleshooting manuals—implicitly defines a graph of
information entities and their relationships. The challenge is to extract
this latent graph structure efficiently with minimal expert effort. We
present a four-stage pipeline.

\placeholder{%
  \textbf{Step 1: Entity extraction.}
  Input: task specification document (e.g., clinical guideline chapter,
  insurance claim form, troubleshooting flowchart).
  Method: Prompt an LLM (e.g., GPT-4o, Qwen-Max) with: ``Given the following
  task specification, extract all required information entities that must be
  collected from the user. For each entity, provide: name, description,
  data type, importance (1--5), and whether it is conditionally required.''
  Aggregate across $k=5$ independent prompts; retain entities appearing in
  $\geq 3/5$ responses.

  \textbf{Step 2: Prerequisite edge identification.}
  For each ordered pair $(u, v)$ of extracted entities, prompt the LLM:
  ``Must entity $u$ be resolved before entity $v$ can be meaningfully
  queried? Answer yes or no and explain why.'' Aggregate across $k=5$
  prompts; edges with $\geq 3/5$ ``yes'' votes are included in $E_P$.
  Verify DAG property (no cycles); if cycles exist, break by removing the
  lowest-confidence edge in each cycle (using Tarjan's algorithm to identify
  strongly connected components).

  \textbf{Step 3: Association edge construction.}
  Embed entity descriptions using BAAI/bge-large. Compute pairwise cosine
  similarity. Retain edges with similarity $\geq \theta$ (default 0.4).
  Alternatively, use LLM-rated semantic relatedness (Likert 1--5; retain
  $\geq 3$). The resulting $G_A$ captures the local dependency structure
  that NCS exploits: entities with high association weight are semantically
  close and changes to one are likely to affect the relevance of the other.

  \textbf{Step 4: Expert validation.}
  Domain expert reviews the auto-generated graph. An important property
  of the NCS principle applies here: because the effect of any single edge
  correction is localized to the immediate neighborhood, experts can
  evaluate and correct edges independently without worrying about cascading
  global effects. This makes the review process modular and efficient.
  Metrics: time to review, number of corrections (edges added/removed/
  modified, nodes added/removed), and post-correction F1 against a
  gold-standard graph if available.
}

\subsection{Evaluation domains}
\label{sec:methods_domains}

\placeholder{%
  \textbf{Domain 1: Hypertension clinical intake.}
  From DrHyper. $|V| = $ [number of entity nodes from the clinical entity
  list; approximately 20--25 top-level groups, expandable to $\sim$60
  individual entities]. Prerequisite graph depth $L = $ [value].
  Association graph density: [value]. Community count: [value].
  Task: systematically collect all required clinical entities for
  hypertension classification, risk stratification, and management planning.
  Ground truth: expert-annotated patient profiles (200 cases from DrHyper;
  58 additional for ablation).
}

\placeholder{%
  \textbf{Domain 2: Diabetes clinical intake.}
  Construct a parallel entity graph for type 2 diabetes initial assessment
  based on ADA Standards of Care. Required entities: HbA1c, fasting glucose,
  BMI, medication history, complication screening (retinal exam, foot exam,
  renal function, cardiovascular risk), lifestyle factors, psychosocial
  screening. $|V| \approx $ [estimate]. $L = $ [estimate].
  Use physician-authored patient profiles for evaluation ([N] cases).
}

\placeholder{%
  \textbf{Domain 3: [Non-medical domain, e.g., insurance claims intake.]}
  Entity graph for auto insurance claim processing: accident details (date,
  location, description), parties involved, injury/damage, police report,
  witness information, policy details, prior claims, liability determination.
  $|V| \approx $ [estimate]. $L = $ [estimate].
  Ground truth from [source: synthetic claim profiles or partnership with
  insurance company].
}

\placeholder{%
  \textbf{Domain 4: [Non-medical domain, e.g., IT technical support triage.]}
  Entity graph for enterprise IT support: symptom description, affected
  systems, error messages, recent changes, environment details (OS, software
  versions), network status, reproducibility steps, business impact,
  workaround status. $|V| \approx $ [estimate]. $L = $ [estimate].
  Ground truth from [source: synthetic support tickets or partnership].
}

\begin{table}[!htbp]
\centering
\caption{Graph structural characteristics across evaluation domains.}
\label{tab:graph_structure}
\small
\input{generated/tab_graph_structure.tex}
\end{table}

\subsection{Baseline implementations}
\label{sec:methods_baselines}

\placeholder{%
  \textbf{Implementation details for each baseline:}

  \textbf{ReAct:} Standard Thought-Action-Observation loop. At each turn,
  the agent generates a thought (``I should ask about X''), asks the
  question, observes the patient response. No persistent memory beyond
  context window. System prompt includes the task description and entity
  list (same information as DualGraph's graph, but as flat text). Critically,
  ReAct performs implicit global reasoning at every turn (the LLM considers
  the entire context window), contrasting with DualGraph's local NCS-based
  updates.

  \textbf{Reflexion:} ReAct + self-reflection. After every $k$ turns
  (default $k=5$), the agent generates a reflection: ``What have I missed?
  What should I prioritize next?'' Reflection is appended to context.
  Reflexion can be viewed as an approximation to global recomputation
  (periodic rather than per-turn), but without graph structure to guide
  the reflection, it lacks the locality guarantees of NCS.

  \textbf{Plan-and-Solve:} At the start of the dialogue, the agent generates
  a complete plan (ordered list of entities to collect). It then executes
  the plan sequentially, marking items as complete. Plan is not revised
  mid-dialogue (tests rigidity). This is the extreme opposite of NCS:
  all computation is front-loaded and no local updates occur during
  execution.

  \textbf{MemGPT / Letta:} Uses the official MemGPT framework with tiered
  memory (main context + archival memory + recall memory). The agent can
  explicitly push/pop/search memory. System prompt includes the task
  description. MemGPT provides memory locality (tiered retrieval) but not
  task-state locality (no graph structure).

  \textbf{Mem0:} Uses the official Mem0 SDK with automatic memory
  management. System prompt includes the task description.

  \textbf{Checklist prompting:} The system prompt includes a structured
  checklist of all required entities with checkboxes. The agent is
  instructed: ``Systematically ask about each unchecked item. Mark items
  as complete when the patient provides the information.'' No graph, no
  dynamic scoring. This is equivalent to DualGraph with $E_P = E_A =
  \emptyset$ and provides a direct test of the value of graph structure.

  \textbf{Memory-only (LLM + a-mem):} Same base LLM with a-mem persistent
  memory. No task control, no checklist. Matches the 58-case baseline from
  DrHyper.

  All baselines use the same base LLM (Qwen-2.5-32B), the same system
  prompt content (task description, entity descriptions), and the same
  patient simulators/profiles. The only difference is the control mechanism.
}

\subsection{Patient / user simulators}
\label{sec:methods_simulators}

\placeholder{%
  \textbf{Clinical domains:}
  Patient simulation by DeepSeek-R1 (different model family from base LLM)
  given a structured patient profile. Same design as DrHyper: no retrieval,
  no guideline access, no knowledge leakage. Profile includes demographics,
  history, lab values, medications, lifestyle, social factors. Simulator
  responds only from profile. Temperature fixed for reproducibility.
  Patient ``character types'' (cooperative, anxious, evasive, low health
  literacy) as in DrHyper.

  \textbf{Non-medical domains:}
  User simulation by the same simulator LLM given a structured case profile
  (e.g., insurance claim details, IT issue description). Same constraints:
  no external knowledge, responds only from profile.

  \textbf{Physician-supervised (clinical pilot):}
  Real patients interact with the system directly; clinician observes in
  real time and can intervene. No simulator involved.
}

\subsection{Evaluation metrics}
\label{sec:methods_metrics}

\placeholder{%
  \textbf{Primary metrics (all domains):}
  \begin{itemize}[noitemsep]
    \item \textbf{Entity acquisition rate}: fraction of required entities
          resolved to confidence $\geq \rho_{\min}$.
    \item \textbf{Information concordance}: accuracy of acquired entity
          values relative to ground-truth profile.
    \item \textbf{Information un-omission}: fraction of required entities
          not omitted (i.e., at least queried, whether or not resolved).
    \item \textbf{Conversation turns}: total turns to termination.
    \item \textbf{Turn efficiency}: acquisition rate divided by turns
          (acquisition per turn).
  \end{itemize}

  \textbf{Secondary metrics:}
  \begin{itemize}[noitemsep]
    \item \textbf{Coherence rating}: human-rated topic coherence (1--5
          Likert) on a sample of [N] dialogues per condition.
    \item \textbf{Prerequisite violation rate}: fraction of turns where an
          entity was queried before all prerequisites were resolved.
    \item \textbf{Redundancy rate}: fraction of turns that re-query an
          already-resolved entity without new information gain.
    \item \textbf{Score stability}: fraction of entity scores unchanged per
          turn (direct measure of NCS effectiveness).
    \item \textbf{CMC} (Conversation Memory Capacity): as defined in DrHyper.
    \item \textbf{IRC} (Information Retrieval Capacity): as defined in
          DrHyper.
  \end{itemize}

  \textbf{Downstream metrics (domain-specific):}
  \begin{itemize}[noitemsep]
    \item Clinical: diagnosis accuracy, risk stratification accuracy,
          medication rationality, lifestyle/follow-up scores.
    \item Insurance: claim adjudication accuracy, missing-info rate.
    \item IT support: resolution rate, escalation accuracy, time to
          resolution.
  \end{itemize}

  \textbf{Statistical analysis:}
  Bootstrap 95\% CIs for all metrics. Welch's $t$-test or Mann--Whitney $U$
  for pairwise comparisons. Bonferroni correction for multiple comparisons
  across domains. Effect sizes (Cohen's $d$). Pre-registration of primary
  endpoints.
}

\subsection{Graph robustness and NCS perturbation analysis}
\label{sec:methods_robustness}

To assess the robustness of the framework to imperfect graph specifications
and to empirically validate the locality predictions of NCS, we conduct
systematic perturbation experiments.

\placeholder{%
  \textbf{Perturbation experiments:}
  \begin{enumerate}[noitemsep]
    \item \textbf{Edge deletion}: Randomly remove $p\%$ of prerequisite edges
          ($p \in \{10, 20, 30, 50\}$) and measure acquisition degradation.
          NCS predicts that deleting an edge $(u, v)$ affects only $v$ and
          its downstream dependents, not the entire graph.
    \item \textbf{Edge addition}: Randomly add $p\%$ spurious prerequisite
          edges and measure turn inflation (unnecessary ordering constraints).
          NCS predicts local effects: the added edge blocks only the target
          node until the spurious prerequisite is resolved.
    \item \textbf{Edge reversal}: Randomly reverse $p\%$ of prerequisite
          edges (wrong ordering) and measure prerequisite violation rate and
          acquisition.
    \item \textbf{Node deletion}: Remove $p\%$ of nodes from the graph
          (simulating incomplete specification) and measure coverage of
          the remaining ground-truth entities.
    \item \textbf{Weight perturbation}: Add Gaussian noise
          $\mathcal{N}(0, \sigma^2)$ to importance weights $w_v$ and measure
          downstream sensitivity. NCS predicts that weight changes affect
          only the perturbed node and its frontier competitors, not the
          global trajectory.
    \item \textbf{NCS validation}: For each perturbation, compare the set of
          nodes whose scores actually changed to the set predicted by NCS
          (the combined neighborhood of perturbed edges/nodes). If NCS is
          valid, these sets should be identical.
  \end{enumerate}
  Run on hypertension domain; verify key findings on one additional domain.
  Report degradation curves (performance vs.\ perturbation rate) and NCS
  prediction accuracy.
}

\subsection{Computational environment}
\label{sec:methods_compute}

\placeholder{%
  Hardware: [GPU type, count, memory].\\
  Software: [Python version, PyTorch version, LLM inference framework].\\
  Inference cost per dialogue: [tokens, time, cost].\\
  Graph construction cost: [time per domain, LLM API calls].\\
  NCS overhead measurement: [describe profiling methodology for measuring
  per-turn score update time with and without NCS].
}


% ============================================================================
% 6. DATA AND CODE AVAILABILITY
% ============================================================================
\section*{Data Availability}

\placeholder{%
  \begin{itemize}[noitemsep]
    \item Task graph specifications (all 4 domains): released on GitHub
          [URL].
    \item Auto-generated and expert-validated graphs: released as JSON on
          GitHub.
    \item Evaluation case profiles (non-clinical domains): released in full.
    \item Clinical case profiles: de-identified data available under data
          access agreement (contact corresponding author).
    \item LoCoMo benchmark: public benchmark [cite].
    \item DualGraph codebase: released on GitHub [URL].
  \end{itemize}
}

\section*{Code Availability}

\placeholder{%
  All code for DualGraph (graph construction, NCS-based node scoring,
  memory module, evaluation harness, baseline implementations) will be
  released at [GitHub URL] under [license]. Evaluation scripts and
  configuration files for all baselines are included. A reference
  implementation of the NCS dirty-flag mechanism is provided as a
  standalone module for integration with other agent frameworks.
}

% ============================================================================
% ACKNOWLEDGEMENTS
% ============================================================================
\section*{Acknowledgements}

\placeholder{Funding sources, institutional support, compute credits.}

% ============================================================================
% AUTHOR CONTRIBUTIONS (NMI requires CRediT)
% ============================================================================
\section*{Author Contributions}

\placeholder{%
  CRediT format.\\
  [Author A]: Conceptualization, Methodology, Software, Writing -- Original
  Draft.\\
  [Author B]: Data Curation, Investigation, Validation.\\
  etc.
}

% ============================================================================
% COMPETING INTERESTS
% ============================================================================
\section*{Competing Interests}

The authors declare no competing interests.


% ============================================================================
% REFERENCES
% ============================================================================

\newpage
\bibliographystyle{unsrtnat}

\begin{thebibliography}{99}

\bibitem{packer2023memgpt}
Packer, C.\ et al.\ MemGPT: Towards LLMs as operating systems.
\textit{arXiv preprint arXiv:2310.08560} (2023).

\bibitem{mem0_2025}
Chhikara, P., Khant, D., Aryan, S., Singh, T.\ \& Yadav, D.\
Mem0: Building production-ready AI agents with scalable long-term memory.
\textit{arXiv preprint arXiv:2504.19413} (2025).

\bibitem{amem_2024}
\placeholder{A-mem reference. Autonomous memory framework for LLM agents.}

\bibitem{yao2023react}
Yao, S.\ et al.\ ReAct: Synergizing reasoning and acting in language models.
\textit{ICLR} (2023).

\bibitem{shinn2023reflexion}
Shinn, N.\ et al.\ Reflexion: Language agents with verbal reinforcement
learning. \textit{NeurIPS} (2023).

\bibitem{wang2023plan}
Wang, L.\ et al.\ Plan-and-Solve prompting: Improving zero-shot
chain-of-thought reasoning by large language models. \textit{ACL} (2023).

\bibitem{traag2019leiden}
Traag, V.A., Waltman, L.\ \& van Eck, N.J.\ From Louvain to Leiden:
guaranteeing well-connected communities. \textit{Scientific Reports}
\textbf{9}, 5233 (2019).

\bibitem{bellman1957dynamic}
Bellman, R.\ \textit{Dynamic Programming}. Princeton University Press (1957).

\bibitem{ref_interaction_failure_1}
\placeholder{Sharma et al.\ (2024). Nature Medicine study on LLM medical
assistants and interaction failure.}

\bibitem{ref_interaction_failure_2}
\placeholder{Additional reference on agent task-completion failures in
deployed systems.}

\bibitem{locomo_ref}
\placeholder{LoCoMo: Long Context and Memory benchmark reference.}

\bibitem{calinescu2011}
Calinescu, G., Chekuri, C., P\'{a}l, M.\ \& Vondr\'{a}k, J.\ Maximizing a
monotone submodular function subject to a matroid constraint. \textit{SIAM
Journal on Computing} \textbf{40}, 1740--1766 (2011).

\bibitem{vondrak2008}
Vondr\'{a}k, J.\ Optimal approximation for the submodular welfare problem in
the value oracle model. \textit{STOC} (2008).

\bibitem{puterman2014mdp}
Puterman, M.L.\ \textit{Markov Decision Processes: Discrete Stochastic
Dynamic Programming}. John Wiley \& Sons (2014).

\bibitem{bertsekas2012dp}
Bertsekas, D.P.\ \textit{Dynamic Programming and Optimal Control}, Vol.~II,
4th ed.\ Athena Scientific (2012).

\bibitem{page1999pagerank}
Page, L., Brin, S., Motwani, R.\ \& Winograd, T.\
The PageRank citation ranking: Bringing order to the web.
\textit{Stanford InfoLab Technical Report} (1999).

\placeholder{%
  \textbf{Additional references needed:}
  \begin{itemize}[noitemsep]
    \item MemGPT/Letta updated reference (2024 version)
    \item Recent agent benchmarks (AgentBench, WebArena, SWE-Bench)
    \item Task-oriented dialogue systems (e.g., MultiWOZ baselines, SGD)
    \item Dialogue state tracking literature
    \item Active information acquisition / adaptive questionnaire literature
    \item Clinical decision support systems reviews
    \item Submodular optimization surveys (Krause \& Golovin, 2014)
    \item Information-theoretic active learning references
    \item Graph-based planning references (HTN planning, STRIPS-like)
    \item Asynchronous dynamic programming references (Bertsekas \&
          Tsitsiklis, 1989)
    \item Value iteration convergence results
    \item Relevant NMI-published agent/LLM papers (2023--2025) for
          positioning
  \end{itemize}
}

\end{thebibliography}


% ============================================================================
% SUPPLEMENTARY INFORMATION
% ============================================================================
\newpage
\appendix
\renewcommand{\thesection}{Supplementary \Alph{section}}

\section{Proof of Theorem~\ref{thm:convergence}}
\label{app:proof}

\placeholder{%
  Full proof of convergence bound. Structure:
  \begin{enumerate}[noitemsep]
    \item Topological layering of $G_P$.
    \item Per-layer coverage analysis using submodularity.
    \item Geometric series bound on expected queries per entity.
    \item Aggregation across layers and entities.
    \item Discussion of tightness and when the bound is loose.
    \item Formal statement and proof that NCS does not affect
          convergence guarantees (NCS changes computation but not
          the mathematical values of scores, so all optimality
          results carry through).
  \end{enumerate}
}

\section{Proof of NCS Properties}
\label{app:ncs_proofs}

\placeholder{%
  \textbf{Proposition~\ref{prop:ncs_holds} extended proof:}
  Full formal proof with edge cases:
  \begin{itemize}[noitemsep]
    \item Handling of the normalization denominator
          $\max_{u \in \mathcal{F}} I(u)$: when the frontier changes
          (an entity is resolved and removed), \emph{all} frontier nodes'
          normalized scores may change. However, this occurs only when a
          frontier node's \emph{own} belief changes (triggering its dirty
          flag) or when a non-frontier node enters the frontier (its
          prerequisite parent was resolved, which is in its neighborhood).
          Formalize this case.
    \item Proof that the dirty-flag implementation is sound (no stale
          scores persist).
    \item Proof that NCS-based computation is equivalent to global
          recomputation (same scores, different computation order).
  \end{itemize}

  \textbf{Dynamic programming connection formalized:}
  \begin{itemize}[noitemsep]
    \item Define the information-gathering MDP formally.
    \item State the Bellman equation (from Remark~\ref{rem:dp}).
    \item Prove the locality lemma: marginal value of querying $v$
          depends only on $\mathcal{N}(v) \cap \mathcal{F}$.
    \item Connect to asynchronous dynamic programming convergence
          results (Bertsekas \& Tsitsiklis, 1989): NCS implements
          a specific asynchronous update schedule determined by the
          graph's edge structure.
    \item Discuss gap between the MDP-optimal policy and the greedy
          NCS policy; bound the approximation ratio.
  \end{itemize}
}

\section{Extended Ablation Results}
\label{app:ablation}

\placeholder{%
  Full ablation tables for all domains.
  Parameter sensitivity plots.
  Graph perturbation degradation curves.
  NCS validation results: comparison of predicted vs.\ actual update sets.
  Score oscillation analysis with and without NCS.
}

\section{Graph Construction: Full Prompts and Examples}
\label{app:graph_prompts}

\placeholder{%
  Complete LLM prompts for entity extraction, prerequisite identification,
  and association scoring.
  Example auto-generated graphs for each domain (with corrections annotated).
  Inter-annotator agreement statistics.
  Analysis of which edge types (prerequisite vs.\ association) are harder
  to auto-generate and why.
}

\section{Domain-Specific Task Graphs}
\label{app:graphs}

\placeholder{%
  Visualizations of the full task graphs for all 4 domains.
  Node and edge lists.
  Community detection results with NCS neighborhood sizes.
  PageRank distributions.
  Neighborhood size distribution (empirical $\Delta$ distribution
  vs.\ $\Delta_{\max}$).
}

\section{Patient / User Simulator Details}
\label{app:simulator}

\placeholder{%
  Full simulator prompts.
  Patient character type definitions and distributions.
  Simulator validation (do simulated patients behave realistically?).
  Oracle cooperation probability $\phi$ calibration by character type.
}

\section{Long-Horizon Memory Benchmark (LoCoMo)}
\label{app:locomo}

To evaluate the memory component independently of the task controller, we
benchmark on LoCoMo~\cite{locomo_ref}, a standardized long-horizon
conversational memory benchmark with single-hop, multi-hop, and temporal
queries over extended multi-session dialogues.

\begin{table}[!htbp]
\centering
\caption{LoCoMo benchmark results. Overall accuracy (\%) and by query
category. Best in bold.}
\label{tab:locomo}
\small
\input{generated/tab_locomo.tex}
\end{table}

\placeholder{%
  \textbf{Analysis.} DualGraph's memory module outperforms because the graph
  controller feeds entity-tagged, confidence-scored, structurally organized
  information into memory rather than raw dialogue text, improving retrieval
  precision. The NCS mechanism further ensures consistent, stable entity
  states that reduce contradictory overwrites, benefiting multi-hop and
  temporal query accuracy. Note: Mem0's published LoCoMo performance
  (reported as $+$26\% over OpenAI Memory~\cite{mem0_2025}) uses a different
  evaluation protocol (their own splits and metrics); our evaluation uses
  the standardized LoCoMo protocol from [cite] for fair comparison across
  all methods.
}

\section{Prospective Pilot: Protocol and Detailed Results}
\label{app:pilot}

\placeholder{%
  IRB approval documentation.
  Consent process.
  Inclusion/exclusion criteria.
  Full results by case.
  Clinician feedback (qualitative and quantitative).
  Safety event log.
}

\section{Downstream Results}
\label{app:downstream}

\placeholder{%
  Full clinical results from DrHyper (200 cases).
  Diabetes downstream results.
  Non-medical domain downstream results.
  Error analysis by risk stratum (from DrHyper Section E.9).
}

\section{Computational Cost Analysis}
\label{app:compute}

\placeholder{%
  Per-turn latency breakdown with and without NCS.
  Total cost per dialogue by domain and method.
  Scaling analysis: cost vs.\ $|V|$ and vs.\ $\Delta_{\max}$.
  NCS speedup factor as a function of graph density.
}


\section{Evolving Inference DAG: Extended Analysis}
\label{app:evolving}

\placeholder{%
  \textbf{A. Emergence detection precision and recall.}\\
  For each domain, report:
  \begin{itemize}[noitemsep]
    \item Precision of emergence detection (fraction of flagged entities
          that are genuinely new vs.\ false positives)
    \item Recall of emergence detection (fraction of ground-truth emergent
          entities correctly detected)
    \item Breakdown by embedding-only filter vs.\ LLM-confirmed filter
          (ablation of the two-stage detection pipeline)
    \item Effect of $\theta_{\mathrm{match}}$ threshold on
          precision-recall tradeoff
  \end{itemize}

  \textbf{B. Graph attachment quality.}\\
  For each emergent node, compare the automatically assigned prerequisite
  and association edges against expert annotation:
  \begin{itemize}[noitemsep]
    \item Edge-level precision and recall
    \item DAG violation rate (proposed edges that would create cycles)
    \item Importance weight correlation with expert ratings
  \end{itemize}

  \textbf{C. NCS containment under graph growth.}\\
  For each node-attachment event across all dialogues:
  \begin{itemize}[noitemsep]
    \item Size of actual update frontier $|\mathcal{U}_t^{\mathrm{attach}}|$
    \item Size of the graph $|V^{(t)}|$ at time of attachment
    \item Ratio $|\mathcal{U}_t^{\mathrm{attach}}| / |V^{(t)}|$
    \item Distribution of these ratios (histogram)
    \item Comparison of NCS-predicted vs.\ actual changed scores (as in
          static-graph NCS validation)
  \end{itemize}

  \textbf{D. Graph growth dynamics.}\\
  Per domain:
  \begin{itemize}[noitemsep]
    \item Distribution of number of emergent entities per dialogue
    \item Distribution of emergence timing (which turn do new entities
          typically appear?)
    \item Final graph sizes $|V^{(\infty)}|$ (distribution)
    \item Types of emergent entities (categorized by domain experts)
  \end{itemize}

  \textbf{E. Failure mode statistics.}\\
  \begin{itemize}[noitemsep]
    \item False-positive emergence rate (duplicates created)
    \item Post-hoc deduplication success rate
    \item Incorrect prerequisite rate (edges later corrected)
    \item Runaway growth cap activations (how often $N_{\max}$ reached)
  \end{itemize}

  \textbf{F. Qualitative examples.}\\
  Full dialogue transcripts (3--4 per domain) showing evolving-DAG
  activation. Annotate with graph snapshots at key turns.
}

\end{document}